# Are accidents caused more by certain demographics? Gender, age, income}

# 

# Are accidents caused more by certain types of cars?

# 

# Are accidents happening more at certain times of the day? Divide day into 3 hr windows maybe

# 

# Are accidents happening more in certain areas of the streets? Intersections, freeways, regular portion of the streets? And what speed limit zones?

# 

# What % are caused due to speeding? Any other reason stands out?

# 

# How many cars on average ( daily /monthly/ annually ) get involved in accidents ( responsible or not )

# 

# Is there any trend where and what times - multiple car ( 3 or more cars ) accidents happen?

# 

# How many accidents are single car accidents? What are the main causes?

```{r}

#install.packages("readr")
library(readr)
library(dplyr)
library(janitor)

getwd()

traffic <- read.csv("Project_Code/Data/data_traffic.csv")


traffic

```

```{r}
## Visualizing some basic data

library(ggplot2)


#By month
ggplot(data = traffic, aes(x=month)) + geom_bar()


#By year
ggplot(data = traffic, aes(x=accident_year)) + geom_bar()


#By gender
pie(table(traffic$party_sex))


#Injury types
pie(table(traffic$collision_severity))


#Time of day

pie(table(traffic$time_cat))


```

```{r}
## This will only focus on who caused accidents

accidenters <- subset(traffic, at_fault == "Yes")

#head(accidenters)

accidenters


#by_age <- subset(accidenters, !is.na(party_age))
#by_age <- subset(by_age, party_age > 15)      

by_age <- filter(accidenters, party_age > 0)

#by_age

##Histogram by age and subdivided by gender
ggplot(data = by_age, aes(x=party_age, fill = factor(party_sex))) + geom_histogram()


```

```{r}


#Compares the types of vehicle
ggplot(traffic, aes(stwd_vehicle_type)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))




primary_streets <- get_dupes(accidenters, primary_rd)
primary_streets 

secondary_streets <- get_dupes(accidenters, secondary_rd)
secondary_streets



important_street <- "MISSION ST"

acc_on_imp <- accidenters

acc_on_imp <- acc_on_imp |>
  subset(primary_rd == important_street) |>
  select(, c(primary_rd, vz_pcf_description))

acc_on_imp


ggplot(acc_on_imp, aes(vz_pcf_description)) + geom_bar()

```
